---
title: "telemetRy: for interacting with telemetry data in R"
output: 
    github_document: default
    html_document: default
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding,
  output_format = "all") })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(telemetRy)
```

## Description
This is a package for interacting with telemetry data using R. Functions for 
analyzing telemetry data over various timescales are provided. These tools were 
developed primarily for working with telemetry datasets generated by Data 
Sciences International ([DSI](https://www.datasci.com)) hardware and exported from Ponemah software.

Note: the author of this package has no affiliation with DSI.

## Installation
telemetRy is available on github. To install telemetRy, you'll need:    
[R > 3.2](https://www.r-project.org)  
[devtools](https://github.com/r-lib/devtoolsdevtools) package from 
Hadley Wickham    
  
First, install devtools:
```{r eval=FALSE}
install.packages('devtools')
```
Then install and load telemetRy:
```{r eval=FALSE}
devtools::install_github(repo = 'andrewstiegler/telemetRy')
library(telemetRy)
```

## Usage
This package comes with a sample dataset generated using a DSI telemetry system.
```{r}
head(sample_BP_data)
```

From this dataset, to calculate telemetry data for a typical day:
```{r}
sample_typical_day <- typical_day(data = sample_BP_data, lights_on = 6)
```

Or to calculate the average data when room lights are off or on:
```{r}
# Output of circadian_avg is a list, select first list element for lights off
sample_lights_off <- circadian_avg(data = sample_BP_data, lights_on = 6)[[1]]
# Select second element for lights on
sample_lights_on <- circadian_avg(data = sample_BP_data, lights_on = 6)[[2]]
```


To see an example of using this package with your own DSI data, check out the 
vignette 
[Importing DSI data](https://andrewstiegler.github.io/telemetRy/DSI_import.html)
