---
title: "Using typical_day on time series data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using typical_day on time series data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(telemetRy)
library(lubridate)
library(ggplot2)
library(data.table)
```

## The typical_day function
The function *typical_day* takes dataframe with a column 
containing time data as input. For a given time of day, 
*typical_days* calculates the average of all other columns in the 
dataframe. This allows recovery of periodic signal even with large amounts of 
noise in the data.

## Example
As an example, we'll first create a dataframe with a time column:
```{r}
df <- data.frame(Time = seq(from = now(), length.out = 200, by = "hours"))
```
Next, we'll create a column of periodic data, with a period of 1 day.
```{r}
# Setting the period to 2*pi/86400 works out to one day because as.duration measures in seconds
df$periodic_data <- 5 * cos((2 * pi / 86400 * 
                               as.numeric(as.duration(df$Time - df$Time[[1]])) + 3)) + 10
```
This periodic function repeats once per day. We'll create a new column for the 
noisy data:
```{r}
set.seed(1234)
df$periodic_data_noisy <- jitter(df$periodic_data, amount = 2)
head(df)
```

Now let's plot the periodic data alone:
```{r fig1, fig.width=5, fig.height=3, fig.align = 'center'}
# Plot the periodic data and connect points to reveal periodicity
ggplot()+
    geom_point(data = df, aes(x = Time, y = periodic_data, color = "Periodic data"))+
    geom_line(data = df, aes(x = Time, y = periodic_data, color = "Periodic curve"))+
    xlab("Time")+
    ylab("Data")+
    scale_color_manual("", values = c("grey60","black"), guide = guide_legend(reverse = TRUE))+
    theme_classic()
```


And then plot the noisy data overlaid with the periodic curve used to generate it:
```{r fig2, fig.width=5, fig.height=3, fig.align = 'center'}
ggplot()+
    geom_point(data = df, aes(x = Time, y = periodic_data_noisy, color = "Noisy data"))+
    geom_line(data = df, aes(x = Time, y = periodic_data, color = "Periodic curve"))+
    xlab("Time")+
    ylab("Data")+
    scale_color_manual("", values = c("black", "grey60"))+
    theme_classic()

```

## Using typical_day

To calculate a typical day of this noisy periodic data:
```{r}
typical_df <- typical_day(data = df, lights_on = 0)
head(typical_df)
```

Output from *typical_day* is a dataframe representing a typical 24-hour period 
from that dataframe. We'll plot the output of *typical_day*:
```{r fig3, fig.height = 3, fig.width = 5, fig.align = 'center'}
ggplot()+
    geom_line(data = typical_df, aes(x = Time/3600, y = periodic_data_noisy, 
                                     color = "Typical noisy data"))+
    xlab("Time (hours)")+
    ylab("Data")+
    scale_color_manual("", values = "blue")+
    theme_classic()
```

## Comparison to non-linear regression model

To determine the parameters of the initial periodic curve from the noisy data 
alone, one approach would be to fit a sinusoidal curve to the data. We can try 
using *nls* for this:

```{r error=TRUE}
# Try NLS fit with reasonable guesses for start
nls_fit <- nls(periodic_data_noisy ~ coef1 * cos(coef2 * as.numeric(as.duration(df$Time - df$Time[[1]]))/86400 + coef3) + coef4, 
               data = df, 
               start = c(coef1 = max(df$periodic_data_noisy) - min(df$periodic_data_noisy), 
                       coef2 = 1, coef3 = 2*pi, coef4 = mean(df$periodic_data_noisy)))

print(coef(nls_fit), digits = 3)
```

Now we'll plot of the nls model fit:
```{r fig4, fig.height = 3, fig.width = 5, fig.align = 'center'}
nls_function <- function (x) {(coef(nls_fit)[1] * cos(coef(nls_fit)[2] * as.numeric(as.duration(x -  df$Time[1]))/86400 + coef(nls_fit)[3]) + coef(nls_fit)[4])
                              }

ggplot()+
  stat_function(fun = nls_function, aes(color = "NLS prediction"))+
  geom_point(data = df, aes(x = Time, y = periodic_data_noisy, color = "Noisy data"))+
  xlab("Time")+
  ylab("Data")+
  scale_color_manual("", values=c("red", "black"))+
  theme_classic()
```
In this case, *nls* finds a local minimum instead of the global minimum and 
cannot fit the periodic trend.

We can try using *nls* on our *typical_day* output instead:
```{r}
nls_typical_fit <- nls(periodic_data_noisy ~ coef1*cos(coef2*as.numeric(typical_df$Time - typical_df$Time[[1]])/86400 + coef3) + coef4, 
               data = typical_df, 
               start = c(coef1 = 1, coef2 = 1, coef3 = 1, coef4 = 1))

print(coef(nls_typical_fit), digits = 3)
```
Note that we didn't need to set starting parameters, which are a common cause of 
instability in *nls* models.


We'll plot the result of running *nls* on the output of *typical_day*:
```{r fig5, fig.height = 3, fig.width = 5, fig.align = 'center'}
nls_typical_function <- function (x) {(coef(nls_typical_fit)[1] * cos(coef(nls_typical_fit)[2]*as.numeric(as.duration(x - df$Time[1]))/86400 
                                                + coef(nls_typical_fit)[3]) +
                                                  coef(nls_typical_fit)[4])
                              }

ggplot()+
  stat_function(fun = nls_typical_function, aes(color = "NLS + typical_day"))+
  geom_point(data = df, aes(x = Time, y = periodic_data_noisy, color = "Noisy data"))+
  xlab("Time")+
  ylab("Data")+
  scale_color_manual("", values=c("blue", "black"))+
  theme_classic()
```

And we'll finally compare the *nls* of *typical_day* to *nls* on the full data:
```{r fig6, fig.height = 3, fig.width = 5, fig.align = 'center', echo=FALSE}
ggplot()+
    stat_function(fun = nls_function, aes(color = "NLS alone"))+
    stat_function(fun = nls_typical_function, aes(color = "NLS + typical_day"))+
    geom_point(data = df, aes(x = Time, y = periodic_data_noisy, color = "Noisy data"))+
    xlab("Time")+
    ylab("Data")+
    scale_color_manual("", values=c("blue", "red", "black"))+
    theme_classic()
```

Using *typical_day* in combination with *nls* reveals periodic data structure 
that *nls* alone misses.

We can also compare the curves alone:
```{r fig7, fig.height = 3, fig.width = 5, fig.align = 'center', echo=FALSE}
ggplot()+
    stat_function(fun = nls_function, aes(color = "NLS alone"))+
    stat_function(fun = nls_typical_function, aes(color = "NLS + typical_day"))+
    geom_line(data = df, aes(x = Time, y = periodic_data, color = "Periodic curve"))+
    xlab("Time")+
    ylab("Data")+
    scale_color_manual("", values=c("blue", "red", "black"))+
    theme_classic()
```
