---
title: "telemetRy: for interacting with telemetry data in R"
output: 
    github_document: default
    html_document: default
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding,
  output_format = "all") })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(telemetRy)
```

## Description
This is a package for interacting with telemetry data using R. Functions for 
analyzing telemetry data over various timescales are provided. These tools were 
developed primarily for working with telemetry datasets generated by Data 
Sciences International ([DSI](https://www.datasci.com)) hardware and exported from Ponemah software.

Note: the author of this package has no affiliation with DSI.

## Installation
telemetRy is available on github. To install telemetRy, you'll need:    
[R > 3.2](https://www.r-project.org)  
[devtools](https://github.com/r-lib/devtoolsdevtools) package from 
Hadley Wickham    
  
First, install devtools:
```{r eval=FALSE}
install.packages('devtools')
```
Then install and load telemetRy:
```{r eval=FALSE}
devtools::install_github(repo = 'andrewstiegler/telemetRy')
library(telemetRy)
```

## Usage
This package comes with a sample dataset generated using a DSI telemetry system.
```{r}
head(sample_BP_data)
```

The package also provides functions for calculating typical days of long-term 
telemetry recordings, and functions for calculating circadian averages of 
telemetry data.

To see an example of using this package with your own DSI data, check out the guide
[Importing DSI data](https://andrewstiegler.github.io/telemetRy/DSI_import.html)

To see an example of the advantages of the *typical_day* function, check out the guide
for [typical_day](https://andrewstiegler.github.io/telemetRy/Typical_avg.html)

